Supporting Information

### 2017-02-11 code/rmarkdown/text.Rmd
address reviewers comments reduce some of the length

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218].

Simply partitioning features based on continuous variables and setting targets for each sub-group may not necessarily result in a solution that secures a representative sample of the continuous variable. For instance, if a feature was split into four sub-features based on a single continuous variable, and a solution captures 10 % of the geographic range of each sub-feature, how much of the original continuous variation does the solution represent? The answer is not immediately clear. If the solution sites reserves evenly along the original variable, then the solution may sample far more than 10 % of the original variation. Alternatively, if the solution sites reserves near the values used to divide the continuous values [as observed when partitioning features based on political boundaries; @r483], then it might sample much less of the original variation. Conventional reserve selection problems cannot differentiate between these solutions--they are both optimal--though the poorer solution may be favored when using additional constraints to reduce cost or fragmentation. This limitation has been known of quite some time [@r218]. In fact, it dates back to some of the earliest reserve selection methods [eg. @r295].

During the early 1990s, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of environmental variation across a study area. Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range. Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482]. However, \texttt{DIVERSITY} has limited utility for species-based conservation planning, in part, because it cannot accommodate multiple biological features--unlike more commonly used tools such as \texttt{Marxan} and \texttt{Zonation}. This fundamental limitation means that \texttt{DIVERSITY} is not useful for most conservation planning scenarios.

@@

# Methods
## PROBLEM FORMULATION
Biodiversity features are defined as the entities that the prioritization is required to preserve (eg. species, ecosystems). Here, attributes are defined as the variation across the feature' range that the prioritization is required to sample. They can be intrinsic variation (eg. genetic or trait) variation that is extrinsic to the feature (eg. environmental conditions). There should, however, be a reasonable underlying hypothesis that relates this variation to the biodiversity processes that the decision maker aims to preserve.

A set of attributes is conceptualized as a space (termed an "attribute space"). For example, a decision maker may require a prioritization that represents populations along climatic gradients. To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this environmental space. Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space. By associating the planning units with climatic data, and calculating a representative statistic for each planning unit (eg. mean), they can be mapped from geographic space to this environmental attribute space. In addition to planning units, attribute spaces also contain demand points.

Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets). The amount of variation in the attribute space that a solution secures depends on the distance between each demand point and each selected planning unit in the attribute space (see below for mathematical formulations). The shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the variation in the spatial attribute. To calculate the proportion of variation secured in a given solution, the distances are scaled by the distance between the demand points to their centroid (equivalent to calculating R$^2$ values for K-means clustering). In any attribute space there may exist points that are impossible (eg. mean annual rainfall -5 mm), or do not occur in the study area (eg. mean annual temperature 30$^{\circ}$C in Antarctica). Additionally, there may be some regions that are desirable for some features and undesirable for others (eg. conditions known to be outside the physiological tolerance of certain species). Thus a different set of demand points and weights may be appropriate for different attribute spaces and features. By siting demand points in desirable regions of an attribute space for a given feature, the decision maker can ensure that prioritizations secure the feature in planning units with spatial attributes that are desirable or representative for that feature.

To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations. However, we can only afford to preserve three of the four populations. To ensure that we preserve a representative sample, we want to preserve populations inhabiting different environmental conditions. We obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional environmental attribute space. Next, we generated demand points as equi-distant points inside this space. By comparing the distributions of demand points and populations in the attribute space, we can identify a suitable prioritization (Fig. 1). We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same environmental characteristics and constitute considerable redundancy if both selected. Instead, a more representative sample of the intra-specific variation could be captured by securing populations $A$ (or $B$), $C$, and $D$. However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, demand points could be sited in conditions that reflect this goal (eg. the top two rows of demand points in Fig. 1). Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized. As we have demonstrated, demand points provide a flexible means to guide the reserve selection process.

The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations. These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451]. Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the unreliable and reliable formulations. The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a solution samples a feature's attribute space. Whereas, the unreliable formulation assumes that features occupy the planning units in which they are found in with 100 % certainty when calculating how well a solution samples a feature's attribute space. For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

@@

Targets are used to ensure that prioritizations adequately preserve each feature. Amount-based targets are used to ensure that the total amount of habitat preserved is sufficient. Let $T_f$ denote the expected amount of area that needs to be preserved for each feature $f \in F$. Space-based targets ensure that a representative or desirable sample of each feature is secured. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. For convenience, all targets are expressed as proportions in the \texttt{R} package.

The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{fs} &= \parbox{25em}{space-based target for feature $f$ in attribute space $s$ according to the corresponding demand points} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM) to penalize fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is assigned to a selected planning unit $J$ where $X_j = 1$. The weighted distance between the demand point and its assigned planning unit $\lambda_{fsi} d_{fsij}$ is used to assess how well the demand point is represented in a given solution. Generally, demand points are represented by their closest selected planning units (unless extremely small space-based targets are used, such as 0.000001, and so the solution effectively unconstrained by the space-based target).

@@

The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is. Constraints (eqn 3b) ensure that all the amount-based targets are met. Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space. For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\lambda_{fsi} d_{fsij} Y_{fsij}$). This total weighted distance is then scaled by the inherent variation in the demand points ($\lambda_{fsi} \delta_{fsi}$). The resulting fraction yields a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis. The constraints ensure that this proportion must be equal to or greater than the space-based target. Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point. Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units. Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.

## Optimization
The unreliable formulation is non-linear. However, the non-linear components can be linearized using existing techniques. The expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al. [-@r426]. The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}). Presently, academics can obtain a license at no cost from the \texttt{Gurobi} website.

@@

To understand and showcase the behavior of the unreliable problem, we conducted a simulation study and two case studies. These studies involved generating solutions using only amount targets to represent prioritizations generated using conventional methods (eg. \texttt{Marxan}), and solutions using amount and space targets using the unreliable formulation. By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection. We have also conducted a benchmark analysis to estimate how long it would take to solve various sized problem (Appendix S2). All analyses were performed in \texttt{R} [version `r getRversion()`; @r192] and all optimization problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across this study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 4a). The second species was simulated to represent a species with an idealized distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the probability density function of a single multivariate normal distribution (eqn 4c). The third species represents a species with two distinct populations (Fig. 2c). This species' distribution was simulated to depict a bimodal distribution, based on the combination of the probability density functions of two multivariate normal distributions (eqn 4d). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 4a--4d. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were set as the planing units' centroids and were weighted according to the units' probability of occupancy.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.1 \tag*{eqn 4a} \\
%
f \left(z, \mu, \Sigma \right) &= \left(2 \pi \right) ^{-1} | \Sigma | e ^{- \frac{1}{2} \left(z - \mu \right)' \Sigma ^{-1} \left(z - \mu \right)} \tag*{eqn 4b} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4c} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
f \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{f \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4d} \\
%
\end{align*}

@@

The uniform species was simulated to occur with a constant probability of occupancy across the study area (Fig. 2a). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount-based targets (Fig. 3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area. This specific configuration is an artifact of the method used to solve the problem. In fact, for this species, all prioritizations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. In the absence of additional criteria, reserve selection methods may or may not return solutions that secure a representative sample of the features. In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points.

The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic spread target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig. 3b). Since all planning units have equal cost and an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf. Fig 3a). Although the use of amount and space based targets has addressed adequacy and representativeness objectives (respectively), they have resulted in a highly fragmented solution.

@@

The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig. 2b). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig. 3e). Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy. Whilst this solution satisfies the adequacy objective for a prioritization in a cost-effective manner--it fails to fulfill the representative objective for a prioritization (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic spread targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is highly fragmented.

@@

The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used. The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this). As with conventional amount-based targets, planners will have to ensure that space targets are high enough to fulfill conservation objectives [eg. when setting targets to conserve intra-specific genetic variation; @r487]. We encourage conservation planners to consult with experts to ensure that appropriate space targets are used.

@@

# Data accessibility
The \texttt{raptr R} package can be downloaded from The Comphrehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}). All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.

>>>>

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain biological features is to conserve a representative sample of the physical attributes that underpin these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity often cannot be divided into operational groups for conservation planning without loss of information [@r489; @r490; @r218]. This limitation has been known of quite some time, and dates back to some of the earliest reserve selection methods [@r489].

To overcome this limitation, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of the environmental variation found across a study area. Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range. Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482]. However, existing \texttt{DIVERSITY}-based methods have limited utility for species-based conservation planning, in part, because they cannot accommodate multiple biological features and do not provide options to specify specific targets pertaining to the amount of habitat required to conserve species (unlike \texttt{Marxan} and \texttt{Zonation}). As a consequence, such methods are not useful for most conservation planning exercises [@r255].

@@

# Methods
## PROBLEM FORMULATION
Biodiversity features are defined as the entities that the prioritization is required to preserve (eg. species, ecosystems). Here, attributes are defined as the variation across the features' ranges that the prioritization is required to sample. These attributes can be intrinsic (eg. genetic or phenotypic) or extrinsic (eg. environmental conditions) to the feature. There should, however, be a reasonable underlying hypothesis that relates the attribute to the biodiversity processes that needs to be conserved.

A set of attributes is conceptualized as a space (termed an "attribute space"). For example, a decision maker may require a prioritization that represents populations along climatic gradients. To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this climatic space. Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space. By associating the planning units with climatic data, and calculating a descriptive statistic for each planning unit (eg. mean), they can be mapped from geographic space to this climatic attribute space. In addition to points representing planning units, attribute spaces also contain demand points.

Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets). For instance, by siting demand points throughout an attribute space, planners can obtain solutions that conserve a representative sample of an attribute space. Alternatively, by siting demand points in specific regions of an attribute space, planners can obtain solutions that target specific samples of an attribute space. For a given set of demand points, the shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the desired variation in the attribute space. In any attribute space there may exist points that are impossible (eg. mean annual rainfall -5 mm), or do not occur in the study area (eg. mean annual temperature 30$^{\circ}$C in Antarctica). Additionally, there may be some regions that are desirable for some features and undesirable for others (eg. conditions known to be outside the physiological tolerance of certain species). Thus a different set of demand points may be appropriate for different attribute spaces and features.

To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations. Since we can only afford to preserve three of the four populations, our objective is to conserve the most representative sample possible. To achieve this goal, we obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional climatic attribute space. Next, we generated demand points as equi-distant points inside this space. By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify the optimal solution (Fig. 1). We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same climatic characteristics and constitutes considerable redundancy if both selected. Instead, a more representative sample of the intra-specific variation could be captured by securing populations _A_ (or _B_), _C_, and _D_. However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, we could site demand points in environmental conditions with temperatures over 30 $^{\circ}$C (ie. the top two rows of demand points in Fig. 1). Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized. Since demand points can be sited and weighted in any configuration, they provide a flexible means to guide the reserve selection process.

The \texttt{raptr R} package utilizes two novel formulations of reserve selection problem to generate prioritizations. These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451]. Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations. The difference between the two formulations is that the reliable formulation explicitly considers the probability that planning units are occupied by features when calculating how well a given solution samples a feature's attribute space. On the other hand, the unreliable formulation assumes that features occupy all of the planning units within which they are found with 100 % certainty when performing these calculations. For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

@@

Targets are used to ensure that prioritizations sufficiently conserve each feature. Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}). Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$. Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved. Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid. This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [@r491; page 106]. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

\begin{align*}
T_s &= \text{amount target for feature $f$} \tag*{eqn 1a}\\
%
\tau_{fs} &= \parbox{25em}{space target for feature $f$ in attribute space $s$} \tag*{eqn 1b}\\
%
B &= \text{boundary length modifier (BLM) to penalize fragmented solutions} \tag*{eqn 1c}\\
\end{align*}

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie. a $j \in J$ where $X_j = 1$). The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$). Generally--unless small space targets are used, eg. 0.000001, so that the solution is effectively unconstrained by the space target--demand points are represented by their closest selected planning units. In poorer quality solutions, demand points will be represented by planning units that have larger distances between them. As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.

@@

The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is. Constraints (eqn 3b) ensure that all the amount-based targets are met. Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space. For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}$). This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$). As previously mentioned, the resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis. The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target. Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point. Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units. Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.

## Optimization
The unreliable formulation is non-linear. However, the non-linear components can be linearized using existing techniques. The expression $X_j X_k$ in (eqn 3a) can be linearized using methods described by Beyer et al. [-@r426]. The \texttt{raptr} R package provides functions to express conservation planning data as optimization problems using linearized versions of the unreliable and reliable formulations (see Appendix S1 for reliable formulation), and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}). Presently, academics can obtain a license at no cost from the \texttt{Gurobi} website.

@@

To understand and showcase the behavior of the unreliable problem, we conducted a simulation study and two case studies. These studies involved generating solutions using only amount targets--equivalent to solutions generated using conventional methods (eg. \texttt{Marxan})--and solutions using amount and space targets using the unreliable formulation. By comparing these solutions, we can gauge the benefits of explicitly including space targets in reserve selection. We have also conducted a benchmark analysis to estimate how long it would take to solve various sized problems (Appendix S2). All analyses were performed in \texttt{R} [version `r getRversion()`; @r192] and all optimization problems were solved to within `r general.params.LST[[MODE]]$MIPGap * 100` % of optimality.

## SIMULATION STUDY
### Methods
We simulated a hypothetical study area with planning units arranged in a $`r sqrt(sim.params.LST[[MODE]]$number.planning.units)` \times `r sqrt(sim.params.LST[[MODE]]$number.planning.units)`$ square grid (Fig. 2). We then simulated three species across the study area. The first species was simulated to represent a hyper-generalist--occurring in all planning units with equal probability (Fig. 2a). This species' distribution was based on a uniform distribution (eqn 4a). The second species was simulated to represent a species with an idealized distribution (Fig. 2b). It has a core range area, and is less likely to be found in areas that are distant from the core area. This species' distribution was simulated using the density function of a single multivariate normal distribution (represented by $\mathcal{N}$; eqn 4b). The third species was simulated to represent a species with two distinct populations (Fig. 2c). It was simulated using the maximum of two density of two multivariate normal distributions (eqn 4c). The probability that each species inhabited a given planning unit was calculated using the $X, Y$ coordinates of the planning unit and eqns 4a--4c. We used a geographic attribute space to showcase the behavior of the problem. For each species, demand points were set as the planing units' centroids and were weighted according to the units' probability of occupancy.

\begin{align*}
\\
P \left( \text{uniformly distributed species} | \left( x, y \right) \right) &= 0.5 \tag*{eqn 4a} \\
%
P \left( \text{normally distributed species} | \left( x , y \right) \right) &= \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 0.0 \\ 0.0 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 12.58&0 \\ 0&12.5 \end{smallmatrix} \right] \right)}{2} \tag*{eqn 4b} \\
%
P \left( \text{bimodally distributed species} | \left( x,y \right) \right) &= \text{Max} \begin{cases}
\matchcal{N} \left(\left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} -3.75 \\ -3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 10&0 \\ 0&10 \end{smallmatrix} \right] \right), \\ \frac{\matchcal{N} \left( \left[ \begin{smallmatrix} x \\ y \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 3.75 \\ 3.75 \end{smallmatrix} \right] \, \left[ \begin{smallmatrix} 8&0 \\ 0&8 \end{smallmatrix} \right] \right)} {2} \\
\end{cases} \tag*{eqn 4c} \\
%
\end{align*}

@@

The uniform species occurred in all planning units with equality probability (Fig. 2a). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount-based targets (Fig. 3a) selected `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units near the southern end of study area. This specific configuration is an artifact of the method used to solve the problem. In fact, for this species, all prioritizations containing `r filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$n` planning units are optimal when considering only `r sim.params.LST[[MODE]]$amount.target*100` % amount targets. In the absence of additional criteria, reserve selection methods may or may not return solutions that secure a representative sample of the features. In this particular case, the solution does not secure a representative sample of the species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % sampled). Note that the proportion here is negative because it performs worse than simply siting a planning unit in the centroid of the species' demand points (equivalent to the $R^2$ of a regression model that explains less variation in the response variable than the mean of the response variable).

The addition of a `r sim.params.LST[[MODE]]$space.target*100` % geographic spread target resulted in a solution that secured a representative sample of the uniform species' range (`r round(filter(sim.spp.DF, species=='Uniform\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % sampled; Fig. 3b). Since all planning units have equal cost and an equal chance of being occupied, this solution has the same number of planning units as the solution generated using only amount targets (cf. Fig 3a). Although the use of amount and space targets has yielded a solution that conserves a sample of the species that contains an adequate amount of habitat and a representative sample of the species' geographic spread, they have yielded a solution that is also highly fragmented solution.

@@

The normally distributed species was simulated to contain a single core area where individuals were most prevalent, and marginal areas where individuals were less likely to occur (Fig. 2b). The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets contained `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$n` planning units and concentrated conservation efforts in the core area (Fig. 3e). Since all planning units have equal costs and areas, this solution contains the planning units with the highest probabilities of occupancy. Whilst this solution conserves a sufficient amount of habitat in a cost-effective manner--it fails to secure a representative sample of the species' geographic spread (`r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount\ntargets')$space.held,2)` % distribution sampled).

The solution generated using `r sim.params.LST[[MODE]]$amount.target*100` % amount targets and `r sim.params.LST[[MODE]]$space.target*100` % geographic spread targets resulted in a solution that is both adequate and representative in terms of the uniform species' distribution (Fig. 3f). This solution used `r filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$n` planning units to secure `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$amount.held,2)` % and sample `r round(filter(sim.spp.DF, species=='Normal\nspecies', prioritisation=='Amount &\nspace targets')$space.held,2)` % of the normal species' distribution. By using amount and geographic spread targets in the reserve selection problem, we obtained a solution that concentrates conservation effort in the range core and also the range margin. However, similar to the corresponding solution for the uniform species (Fig. 3b), this solution is also highly fragmented.

@@

The degree to which a prioritization truly secures a representative sample of a feature depends on (1) the attribute space(s), (2) the distribution of demand points chosen by the conservation planner, and (3) the space target used. The optimal solution will not effectively represent the decision maker's objectives if an inappropriate set of spatial variables or demand points are used to represent the attribute space. Demand points should be distributed across the full range of variation in an attribute space to obtain solutions that secure a representative sample of the attribute space (see the `make.DemandPoints` function for statistical routines to achieve this). As with conventional amount-based targets, planners will have to ensure that space targets are high enough to fulfill conservation objectives [eg. when setting targets to conserve intra-specific genetic variation; @r487]. We encourage conservation planners to consult with experts to identify suitable targets.

@@

# Data accessibility
The \texttt{raptr R} package can be downloaded from The Comprehensive R Archive Network (\url{https://CRAN.R-project.org/package=raptr}). All data, code, and results are stored in an online repository (\url{https://github.com/jeffreyhanson/raptr-manuscript}) to permit replication and validation of this study.
